\section{Conclusion}
\label{sec:conclusion}

We presented \fedsink{}, a Byzantine-resilient federated governance protocol for
multi-agent AI systems.  The protocol combines spectral-sphere projection with residual
injection (preventing Birkhoff Uniformity Collapse \cite{mcfs_iclr2027}), Gaussian
$(\varepsilon,\delta)$-differential privacy (preventing inference of private agent state),
zk-SNARK MACI proofs (preventing bribery attacks on routing decisions), and Merkle-CRDT
gossip (achieving eventual consistency under $f < N/3$ Byzantine faults).

Our main security results are:
\begin{itemize}
  \item \textbf{CID Integrity} (Theorem~\ref{thm:cid_integrity}): Byzantine nodes cannot
    corrupt honest replicas without detection.
  \item \textbf{EEC} (Theorem~\ref{thm:eec}): honest replicas converge to identical state
    in $O(\log N)$ gossip rounds.
  \item \textbf{$(\varepsilon,\delta)$-DP} (Theorem~\ref{thm:dp_guarantee}): agent trust
    weights are differentially private.
\end{itemize}

A structural observation (Lemma~\ref{lem:sensitivity}) is that residual injection reduces
DP noise requirements by $(1-\alpha)$ multiplicatively — at $\alpha{=}0.1$, a $10\%$
reduction with no tradeoff in stability or security.  The stability fix and the
cryptographic efficiency improvement are synergistic.

\paragraph{Limitations.}
(1) The zk-SNARK circuit for Gaussian randomness verification requires a VRF or
hash-to-curve, adding significant proving overhead — benchmarking is deferred to Phase 3.
(2) Official SWE-bench evaluation is not claimed here; routing evidence is limited to
the local synthetic SWE-bench-shaped evaluator.
(3) Timing side-channels and traffic analysis are out of scope for v1.
(4) Advanced composition accounting (moments accountant) for multi-round DP is not yet
integrated.

\paragraph{Future work.}
\begin{itemize}
  \item \textbf{Tight sensitivity via BODES composition.}  When $H_{\mathrm{proj}}$
    derives from BODES latent steering vectors, the output space may be a strict subset
    of $\mathcal{S}_r$, lowering $\Delta g$ further below $2(1-\alpha)r$.

  \item \textbf{Adaptive $\varepsilon$ scheduling.}  High-stakes rounds (constitutional
    amendments) should use tight $\varepsilon$; routine routing can use loose $\varepsilon$.
    A session-level privacy account with adaptive per-round allocation is unexplored.

  \item \textbf{Efficient zk-SNARK circuits.}  The Gaussian draw verification circuit
    is the bottleneck.  Plonky2 or STARKs may offer better proving times at the cost
    of larger proof sizes.

  \item \textbf{Threshold decryption for trust aggregation.}  Rather than each agent
    broadcasting a full matrix, threshold FHE could allow the swarm to compute aggregate
    trust updates without any single agent learning individual contributions.
\end{itemize}

\paragraph{Reproducibility.}
All protocol steps are implemented in the \texttt{constitutional\_swarm} package
(constitutional hash \texttt{608508a9bd224290}).  Phase 2 regression tests cover CID
integrity, EEC convergence, and DP calibration, and
\texttt{scripts/reproduce\_paper\_claims.py} reproduces the local gossip, Byzantine
rejection, synthetic routing, and latency-harness checks.  Official external SWE-bench
results remain outside the reported claims.  Per-metric governance evolution is recorded
via the
\texttt{EvolutionLog} component, which enforces strict monotonicity and strict
acceleration on tracked metrics at write time, providing a structural audit trail
complementary to the cryptographic guarantees of \fedsink{}.
