\section{Evaluation}
\label{sec:evaluation}

We evaluate \fedsink{} on three dimensions:
\textbf{(E1)} Protocol correctness and convergence under Byzantine faults.
\textbf{(E2)} DP noise accuracy and sensitivity tightening.
\textbf{(E3)} local synthetic SWE-bench-shaped routing behavior.  Official SWE-bench
results are not claimed in this paper.

\subsection{Setup}

\paragraph{Implementation.}
\fedsink{} is implemented in the \texttt{constitutional\_swarm} package
(constitutional hash \texttt{608508a9bd224290}).
Spectral projection uses \texttt{torch.linalg.svd}; Gaussian noise uses
\texttt{torch.randn}; CID uses SHA3-256.  zk-SNARK proofs are simulated (full
Groth16 circuit benchmarking is a Phase 3 deliverable).

\paragraph{Network topology.}
The local reproducibility harness evaluates Merkle-CRDT gossip convergence with
deterministic pseudo-random peer selection and configurable agent counts, artifact
rates, and gossip fanout.  It also runs a deterministic Byzantine tamper-injection
check that attempts to merge invalid content-addressed nodes into an honest replica.

\subsection{Protocol Correctness (E1)}

\begin{table}[t]
  \centering
  \footnotesize
  \caption{Protocol correctness benchmark plan under Byzantine faults.
    Local reproducibility artifacts are listed in docs/internal/claims\_map.md.}
  \label{tab:protocol_correctness}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{@{}p{0.10\linewidth}p{0.12\linewidth}p{0.23\linewidth}p{0.21\linewidth}p{0.09\linewidth}@{}}
    \toprule
    $N$ & $f$ & Accept. & Rounds & EEC \\
    \midrule
    20 & $<N/3$ & tampered CIDs rejected & converged by final merge & Yes \\
    config. & $<N/3$ & JSON & JSON & JSON \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:protocol_correctness} records the local benchmark contract rather than
machine-specific timing values.  Existing regression tests and
\texttt{scripts/reproduce\_paper\_claims.py} cover honest-replica convergence and
Byzantine exclusion.

\paragraph{CID integrity verification.}
Current Merkle-CRDT tests reject tampered tuples by content hash, and the local harness
reports attempted versus accepted tampered nodes as JSON.

\subsection{Differential Privacy Accuracy (E2)}

\begin{table}[t]
  \centering
  \caption{Empirical noise $\sigma$ vs.\ theoretical calibration.
    $r=1.0$, $\delta=10^{-5}$, $\alpha=0.1$.
    ``Relative error'' is $|{\sigma_{\mathrm{emp}} - \sigma_{\mathrm{theory}}}|/\sigma_{\mathrm{theory}}$.}
  \label{tab:dp_accuracy}
  \begin{tabular}{ccccc}
    \toprule
    $\varepsilon$ & $\sigma_{\mathrm{theory}}$ & $\sigma_{\mathrm{emp}}$ & Rel.\ error & DP verified \\
    \midrule
    1.0 & 8.721 & formula check & exact & Yes \\
    2.0 & 4.360 & formula check & exact & Yes \\
    4.0 & 2.180 & formula check & exact & Yes \\
    8.0 & 1.090 & formula check & exact & Yes \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:dp_accuracy} confirms that implementation calibration matches
equation~\eqref{eq:sigma}.  The $10\%$ noise reduction from $\alpha{=}0.1$ residual
injection (Lemma~\ref{lem:sensitivity}) is verified: at $\varepsilon{=}2.0$, baseline
$\sigma = 4.84$ vs.\ residual-injection
$\sigma = 4.36$, a reduction of $10\%$.

\subsection{SWE-bench-Shaped Routing Evaluation (E3)}
\label{sec:evaluation_sweb}

\textbf{Official SWE-bench results are not claimed.}  This section reports only the
local synthetic evaluator that exercises the same routing surface without Docker,
network access, or external benchmark images.

\paragraph{Experimental setup.}
$N$ agents are deployed on SWE-bench-shaped coding tasks with deterministic hidden
domain competencies.  Agents are specialized by domain and trust-weighted via the
SpectralSphere matrix.  Task routing is determined by the projected trust matrix,
matching the routing interface used by the package SWE-bench scaffold.

\paragraph{Baselines.}
\begin{itemize}
  \item \textbf{Flat routing}: all agents receive all tasks with uniform probability.
  \item \textbf{Centralized trust}: oracle trust matrix (upper bound, not privacy-preserving).
  \item \textbf{Sinkhorn-CRDT}: same protocol as \fedsink{} but Sinkhorn projection
    instead of SpectralSphere (expected to collapse, Section~\ref{sec:protocol}).
  \item \textbf{\fedsink{} (ours)}: full protocol as described.
\end{itemize}

\paragraph{Metrics.}
Task success rate (fraction of SWE-bench instances resolved), routing diversity
(variance of task assignments across agents), and convergence overhead (gossip rounds
per task assignment).
All per-epoch metrics will be recorded via the \texttt{EvolutionLog} component, which
enforces strict monotonicity and strict acceleration on governance metric series at
write time: a regression or deceleration in task success rate is rejected structurally
before being persisted, turning the audit log into a safety contract rather than a
post-hoc record.

\paragraph{Local outcome.}
The local evaluator reports per-seed resolve rates for trust-weighted routing versus a
round-robin baseline and emits a mean lift score.  The JSON payload sets
\texttt{official\_swe\_bench\_claimed=false} so synthetic routing evidence cannot be
mistaken for official benchmark performance.

\begin{table}[b]
  \centering
  \footnotesize
  \caption{Local SWE-bench-shaped evaluator contract. Official SWE-bench remains
    outside the reported results.}
  \label{tab:sweb_results}
  \begin{tabular}{p{0.30\linewidth}p{0.13\linewidth}p{0.22\linewidth}p{0.14\linewidth}}
    \toprule
    Method & Success & Routing div. & Conv. \\
    \midrule
    Round-robin routing    & JSON & N/A & N/A \\
    \fedsink{} synthetic   & JSON & JSON & local \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Computational Overhead}

\begin{table}[t]
  \centering
  \caption{Protocol step latency harness per agent per round ($n{=}50$, CPU).
    Values are emitted as machine-specific JSON rather than fixed paper constants.}
  \label{tab:latency}
  \footnotesize
  \begin{tabular}{p{0.55\linewidth}p{0.28\linewidth}}
    \toprule
    Step & Latency (ms) \\
    \midrule
    Spectral projection (power iteration, $n{=}50$) & JSON ns/op \\
    Residual injection & JSON ns/op \\
    DP calibration  & formula check \\
    CID append & JSON ns/op \\
    zk-SNARK proof (simulated) & TBD (Phase 3) \\
    CRDT merge + gossip & JSON convergence \\
    \midrule
    Total (excl.\ zk-SNARK) & not fixed across CPUs \\
    \bottomrule
  \end{tabular}
\end{table}

The dominant expected cost (excluding the zk-SNARK, which requires full circuit
benchmarking in Phase 3) is the SVD, which scales as $O(n^3)$.  Latency microbenchmarks
are reproducible through the script, but we do not report CPU-specific millisecond
constants or task-overhead percentages as portable claims.
