\section{Introduction}
\label{sec:introduction}

The deployment of multi-agent AI systems for high-stakes tasks —
software engineering, medical triage, legal analysis, and autonomous governance — raises
a fundamental security question: how do agents maintain a consistent, tamper-evident
shared state while broadcasting sensitive information over an open peer-to-peer network?

\paragraph{The governance broadcast problem.}
In constitutional multi-agent systems \cite{bai2022constitutional,acgs2025}, each agent
maintains an internal \emph{trust-routing matrix} $H \in \R^{n \times n}$ encoding its
confidence in peers' specialized capabilities.  Agents must broadcast updates to $H$
so that the swarm maintains globally coherent routing.  But broadcasting $H$ in the
clear exposes two attack surfaces:

\begin{itemize}
  \item \textbf{Inference attack (honest-but-curious nodes).}  An adversary observing
    the broadcast stream can reverse-engineer an agent's latent activation space,
    private prompts, or constitutional trust model from the routing matrix entries.

  \item \textbf{Sybil routing attack (Byzantine nodes).}  A Byzantine node $\mathcal{B}$
    can inject forged trust updates that redirect high-value tasks toward
    $\mathcal{B}$-controlled executors, violating Multi-Agent Constitutional Integrity
    (MACI) \cite{maci_spec}.  This is structurally a bribery attack: $\mathcal{B}$
    ``buys'' routing priority by manipulating the shared trust state.
\end{itemize}

\paragraph{Our approach: \fedsink{}.}
We present \fedsink{} (\textbf{Fed}erated \textbf{Sink}horn with \textbf{M}ulti-\textbf{A}gent
\textbf{C}onstitutional \textbf{I}ntegrity), a protocol that addresses both attack surfaces
through a layered defense:

\begin{enumerate}
  \item \textbf{Spectral projection + residual injection} ensures broadcast matrices
    satisfy $\spnorm{H} \leq r$ and prevents Birkhoff Uniformity Collapse
    \cite{mcfs_iclr2027}, maintaining diverse trust routing even under adversarial
    perturbation.

  \item \textbf{Gaussian $\varepsilon$-DP noise} added to the projected matrix before
    broadcast provides $({\varepsilon}, \delta)$-differential privacy
    \cite{dwork2014algorithmic}, preventing inference of private state.

  \item \textbf{zk-SNARK proof (MACI)} \cite{maci_spec} asserts that the broadcast matrix
    was computed correctly from a valid local state and that the spectral constraint was
    enforced.  An agent cannot prove to a colluder that it routed a specific task in a
    specific direction, eliminating bribery incentives.

  \item \textbf{Merkle-CRDT gossip} \cite{kleppmann2022crdt} provides Byzantine-resilient
    eventual consistency: honest replicas accept only CID-verified, proof-validated updates,
    converging to identical state in $O(\log N)$ rounds.
\end{enumerate}

\paragraph{Key insight: residual injection tightens DP.}
A structural observation (Lemma~\ref{lem:sensitivity}) is that residual injection —
introduced in \cite{mcfs_iclr2027} to prevent trust collapse — also reduces the
$\ell_2$ sensitivity of the broadcast function from $2r$ to $2(1-\alpha)r$.  At
$\alpha = 0.1$, this reduces required Gaussian noise by $10\%$ for the same
$(\varepsilon, \delta)$ budget.  Stability and privacy are synergistic rather than
competing.

\paragraph{Contributions.}
\begin{itemize}
  \item \fedsink{} protocol specification
    (Section~\ref{sec:protocol}), combining spectral projection, residual injection,
    $\varepsilon$-DP, zk-SNARK MACI, and Merkle-CRDT gossip in a single coherent
    seven-step protocol.

  \item Formal security proofs (Section~\ref{sec:security}): CID Integrity
    (Theorem~\ref{thm:cid_integrity}), Eventual Epistemic Convergence
    (Theorem~\ref{thm:eec}), and $\varepsilon$-DP guarantee
    (Theorem~\ref{thm:dp_guarantee}).

  \item Sensitivity tightening via residual injection (Lemma~\ref{lem:sensitivity}),
    reducing DP noise by $(1-\alpha)$ multiplicatively.

  \item Empirical evaluation on synthetic swarm benchmarks and SWE-bench coding tasks
    (Section~\ref{sec:evaluation}; SWE-bench results pending Phase 3/4 completion).
\end{itemize}

\paragraph{Scope.}
\fedsink{} v1 is out of scope for: timing side-channels, traffic analysis, compromised
VRF oracles, and adaptive $\varepsilon$ scheduling.  These are directions for future
work (Section~\ref{sec:conclusion}).
