\section{The \fedsink{} Protocol}
\label{sec:protocol}

\subsection{System Model}

Let $\mathcal{N} = \{1, \ldots, N\}$ be a set of $N$ agents communicating over an
asynchronous peer-to-peer network.  Each agent $i \in \mathcal{N}$ has:
\begin{itemize}
  \item A local dataset $D_i$ (activations, task history, capability embeddings).
  \item A local trust computation function $\phi_i : D_i \to \R^{n \times n}$ producing
    a raw routing matrix $H_{\mathrm{new}}$.
  \item A VRF key pair $(\mathsf{sk}_i, \mathsf{vk}_i)$ for verifiable randomness.
  \item A zk-SNARK proving key $\mathsf{pk}$ (shared, from a trusted setup).
\end{itemize}

Up to $f = \lfloor N/3 \rfloor$ agents may be \emph{Byzantine}: they deviate
arbitrarily from the protocol.  Remaining agents are \emph{honest}: they follow the
protocol specification exactly.

\subsection{Threat Model}

\paragraph{Honest-but-curious peers.}
These agents read Merkle-CRDT entries to update their local governance state and attempt
to infer private data (activations, prompts, constitutional threat models) of other agents
from the broadcast routing matrices.  Defense: Gaussian $\varepsilon$-DP noise.

\paragraph{Byzantine / colluding nodes.}
These agents attempt to inject forged trust updates to redirect high-value tasks toward
controlled executors (Sybil routing attack) or to corrupt the shared CRDT state.
Defense: zk-SNARK proof verification + CID integrity checking.

\paragraph{Out of scope (v1).}
Timing side-channels, traffic analysis, compromised VRF oracles.

\subsection{Cryptographic Primitives}

\begin{definition}[Content Identifier]
  The content identifier $\MCid(\tilde{H}) = \mathsf{H}_{\mathrm{CR}}(\tilde{H})$ is
  computed by a collision-resistant hash function $\mathsf{H}_{\mathrm{CR}}$ (e.g., SHA3-256).
\end{definition}

\begin{definition}[Merkle-CRDT]
  A Merkle-CRDT is a conflict-free replicated data type (CRDT) \cite{shapiro2011crdt}
  augmented with a Merkle-DAG structure: each state update is a content-addressed block
  referencing its causal predecessors.  The merge function is commutative, associative,
  and idempotent (CAI).  Byzantine fault tolerance is achieved by requiring valid
  zk-SNARK proofs for all accepted entries \cite{kleppmann2022crdt}.
\end{definition}

\begin{definition}[zk-SNARK for \fedsink{}]
  \label{def:zksnark_circuit}
  The zk-SNARK proof $\pi$ for a broadcast tuple $(\MCid, \tilde{H}, \pi)$ asserts
  the following circuit statement:
  \[
    \exists\, H_{\mathrm{proj}},\, Z \;\text{ s.t. }\;
    \begin{cases}
      \spnorm{H_{\mathrm{proj}}} \leq r \\
      \tilde{H} = (1-\alpha)H_{\mathrm{proj}} + \alpha I + Z \\
      Z \sim \mathcal{N}(0, \sigma^2 I) \text{ via VRF}
    \end{cases}
  \]
  Soundness: no PPT prover can produce an accepting proof for a false statement
  (except with negligible probability).
  Zero-knowledge: the proof reveals nothing about $H_{\mathrm{proj}}$ or $Z$ beyond
  the circuit statement.
\end{definition}

\subsection{Seven-Step Protocol Execution}

The \fedsink{} protocol consists of seven steps executed by each honest agent $i$
at each governance round.

\paragraph{Step 1 — Local trust computation.}
Agent $i$ computes $H_{\mathrm{new}} \leftarrow \phi_i(D_i)$ from local data.
$H_{\mathrm{new}}$ is never broadcast.

\paragraph{Step 2 — Spectral projection.}
\begin{equation}
  \label{eq:spectral_proj}
  H_{\mathrm{proj}} \leftarrow \Pi_r(H_{\mathrm{new}}) = U\,\mathrm{diag}(\min(\sigma_j, r))\,V^\top,
\end{equation}
where $H_{\mathrm{new}} = U\Sigma V^\top$ is the SVD.  This enforces
$\spnorm{H_{\mathrm{proj}}} \leq r$ and prevents Birkhoff Uniformity Collapse
\cite{mcfs_iclr2027}.

\paragraph{Step 3 — Residual injection.}
\begin{equation}
  \label{eq:residual}
  H_{\mathrm{proj}} \leftarrow (1-\alpha)\,H_{\mathrm{proj}} + \alpha I,
  \quad \alpha = 0.1.
\end{equation}
This provides topological stability (Section~\ref{sec:security}) and tightens the DP
sensitivity by factor $(1-\alpha)$ (Lemma~\ref{lem:sensitivity}).

\paragraph{Step 4 — DP noise addition.}
Calibrate $\sigma$ to Lemma~\ref{lem:sensitivity}:
\begin{equation}
  \label{eq:sigma}
  \sigma = \frac{2(1-\alpha)r \cdot \sqrt{2\ln(1.25/\delta)}}{\varepsilon}.
\end{equation}
Sample $Z \sim \mathcal{N}(0, \sigma^2 I_{n^2})$ using VRF-seeded randomness and
compute the noisy matrix:
\begin{equation}
  \label{eq:noisy_matrix}
  \tilde{H} \leftarrow H_{\mathrm{proj}} + Z.
\end{equation}

\paragraph{Step 5 — zk-SNARK proof generation.}
Agent $i$ generates proof $\pi \leftarrow \mathsf{Prove}(\mathsf{pk},\,
H_{\mathrm{new}},\, H_{\mathrm{proj}},\, Z,\, \mathsf{vk}_i)$ asserting the circuit
statement of Definition~\ref{def:zksnark_circuit}.  The proof is non-interactive and
publicly verifiable.

\paragraph{Step 6 — Merkle-CRDT broadcast.}
Compute $\MCid \leftarrow \mathsf{CID}(\tilde{H})$ and append the tuple
$(\MCid,\, \tilde{H},\, \pi)$ to the Merkle-CRDT.  The CRDT records the causal
predecessors of this update (Merkle-DAG links).

\paragraph{Step 7 — Peer verification and gossip.}
When agent $j \neq i$ receives $(\MCid,\, \tilde{H},\, \pi)$:
\begin{enumerate}
  \item Verify $\MCid = \mathsf{CID}(\tilde{H})$.  If false, \textbf{reject}.
  \item Verify $\mathsf{Verify}(\mathsf{vk},\, \tilde{H},\, \pi) = 1$.
    If false, \textbf{reject}.
  \item If both checks pass, \textbf{accept}: merge into local CRDT state and
    gossip to peers.
\end{enumerate}
Invalid proofs are never forwarded (no gossip amplification of Byzantine messages).

\subsection{Privacy Composition under Multiple Rounds}

If agent $i$ broadcasts $k$ updates per session, the composed privacy cost under
standard composition is approximately $(\varepsilon\sqrt{2k\ln(1/\delta)},\, k\delta)$.
Under advanced composition (the moments accountant \cite{abadi2016moments}), tighter
budgets are achievable.  We recommend allocating $\varepsilon_{\mathrm{session}}$ at
session start and tracking consumption per round.

\subsection{Impact on Swarm ODE Dynamics}

Because $Z$ has zero mean ($\mathbb{E}[Z] = 0$), the macroscopic mean-field dynamics
are unaffected:
\begin{equation}
  \frac{d\bar{H}}{dt} = f_\theta(\bar{H},\, t) + \mathbb{E}[Z(t)] = f_\theta(\bar{H},\, t).
\end{equation}
The noise term acts as Langevin dynamics: at high $\sigma$ (tight privacy), the swarm
explores the trust manifold stochastically; at low $\sigma$ (loose privacy), it converges
deterministically.  For matrix Gaussian noise, the leading spectral-norm scale is
$2\sigma\sqrt{n}$; the conservative regime $\sigma \leq r/(2\sqrt{n})$ keeps the expected
perturbation within the same scale as the spectral constraint.
