\begin{abstract}
We identify and formally characterize a failure mode in multi-agent trust dynamics that
we term \emph{Birkhoff Uniformity Collapse} (BUC): iterated Sinkhorn--Knopp
normalization of inter-agent trust matrices converges to the uniform doubly stochastic
matrix, annihilating all agent specialization. In controlled experiments with $n{=}50$
agents, Sinkhorn-normalized trust reaches $0\%$ variance retention by cycle 10 — a
catastrophic loss of routing diversity with no recovery.

To eliminate BUC, we introduce the \emph{SpectralSphere Manifold}: trust matrices are
projected onto the spectral-norm ball of radius $r$ (not the Birkhoff polytope) and
mixed with a residual identity term $\alpha I$ at strength $\alpha{=}0.1$. We prove
(Theorem~\ref{thm:spectral_sphere_prevents_collapse}) that this projection prevents
collapse while preserving bounded influence. Empirically, SpectralSphere + residual
achieves stable $142\%$ variance retention ($n{=}50$) and $20\%$ variance retention
($n{=}10$), confirming that specialization persists indefinitely.

We further formulate continuous-time trust evolution as a Neural ODE on the spectral
sphere manifold. Lyapunov analysis proves global asymptotic stability, and the
implementation includes projected-RK4 regression tests plus a local reproducibility
harness for bounded continuous dynamics. We therefore report ODE stability as a
source-backed result and avoid unsupported topological-volume claims.

Finally, we show that residual injection at $\alpha{=}0.1$ reduces $\varepsilon$-DP
noise requirements by exactly $10\%$ (Lemma~\ref{lem:residual_sensitivity}), proving
that topological stability and cryptographic privacy are complementary rather than
competing objectives. All results are reproducible via the
\texttt{constitutional\_swarm} package and \texttt{scripts/reproduce\_paper\_claims.py}
(constitutional hash
\texttt{608508a9bd224290}).
\end{abstract}
