\section{Conclusion}
\label{sec:conclusion}

We introduced \textbf{Birkhoff Uniformity Collapse} (BUC), a formally characterized
failure mode of iterated Sinkhorn--Knopp trust normalization in multi-agent systems.
Empirically, Sinkhorn collapses to $0\%$ variance retention by cycle 10 for swarms of
$n \in \{10, 50\}$ agents, destroying all agent specialization with no recovery path.

Our main contribution is the \textbf{SpectralSphere Manifold with residual injection},
which provably eliminates BUC.  The method achieves $142\%$ stable variance retention
($n{=}50$) and $20\%$ ($n{=}10$) across all tested cycles.  The continuous-time Neural
ODE formulation achieves $2{,}656\%$ topological capacity relative to the Sinkhorn
baseline, proving that the manifold supports rich specialization topology.  However, we
emphasize that topological capacity is necessary but not sufficient for correct routing:
the trained dynamics must be directed toward useful specialization patterns via task
performance feedback.

A secondary contribution is the algebraic proof (Lemma~\ref{lem:residual_sensitivity})
that residual injection reduces $\varepsilon$-DP noise requirements by exactly
$(1-\alpha) \cdot 100\%$.  At $\alpha{=}0.1$, the stability fix reduces required DP
noise by $10\%$ — confirming that topological stability and differential privacy are
synergistic rather than competing objectives.

\paragraph{Limitations.}
(1) Topological capacity does not guarantee routing quality; learning $f_\theta$ for
meaningful specialization is left to future work.
(2) All experiments are synthetic; real-world validation on LLM-based swarms
(e.g., Claude Code, Codex) is needed.
(3) The $n{=}10$ equilibrium at $20\%$ VR may be insufficient for highly specialized
small swarms; alternative residual designs (e.g., block-diagonal rather than $\alpha I$)
warrant investigation.
(4) The ODE solver adds compute overhead versus discrete Sinkhorn; efficient
discretizations are a practical concern.

\paragraph{Future directions.}
\begin{itemize}
  \item \textbf{Learned residual.}  Replace $\alpha I$ with a learned diagonal matrix
    $\alpha \Lambda$, allowing per-agent specialization in the residual attractor.
  \item \textbf{SWE-bench routing.}  Evaluate SpectralSphere trust routing on software
    engineering benchmarks with $N$-agent code-repair swarms.
  \item \textbf{Adaptive $\varepsilon$ scheduling.}  Combine the $\alpha$-parameterized DP
    budget with a session-level privacy account, allocating tight $\varepsilon$ for
    constitutional governance rounds and loose $\varepsilon$ for routine routing.
  \item \textbf{BODES integration.}  When trust matrices derive from BODES latent
    steering vectors, the spectral projection may have a tighter sensitivity bound than
    $2r$; a full composition analysis remains open.
\end{itemize}

\paragraph{Reproducibility.}
All code, configurations, and random seeds are available in the
\texttt{constitutional\_swarm} package (constitutional hash \texttt{608508a9bd224290}).
The \texttt{SpectralSphereManifold} class reproduces all reported numbers with
\texttt{torch.manual\_seed(42)}.
