\section{Method}
\label{sec:method}

\subsection{Setup and Notation}

Let $\mathcal{A} = \{1, \ldots, n\}$ be a swarm of $n$ agents.  The inter-agent trust
state at discrete time $k$ is a matrix $H^{(k)} \in \R^{n \times n}$, where
$H^{(k)}_{ij}$ encodes agent $i$'s confidence in agent $j$'s competence on the current
task distribution.  We write $\spnorm{\cdot}$ for the spectral norm (largest singular
value), $\|\cdot\|_F$ for the Frobenius norm, and $\mathrm{Var}(H)$ for the variance of
the eigenvalue distribution of $H$.

\subsection{Birkhoff Uniformity Collapse}

\begin{definition}[Sinkhorn Update Rule]
  \label{def:sinkhorn_rule}
  The Sinkhorn update applies $t$ iterations of alternating row and column normalization
  to a positive matrix $H > 0$:
  \[
    \mathrm{Sink}_t(H) := \lim_{t' \to t} \left(D_r^{(t')} H D_c^{(t')}\right),
  \]
  where $D_r^{(t')}, D_c^{(t')}$ are diagonal scaling matrices enforcing
  $H\mathbf{1} = \mathbf{1}$ and $H^\top\mathbf{1} = \mathbf{1}$ alternately.
\end{definition}

\begin{definition}[Birkhoff Uniformity Collapse]
  \label{def:buc}
  A trust dynamics sequence $\{H^{(k)}\}_{k \geq 0}$ exhibits \textbf{Birkhoff
  Uniformity Collapse} (BUC) if
  \[
    \lim_{k \to \infty} \mathrm{Var}\!\left(H^{(k)}\right) = 0
    \quad \text{and} \quad
    \lim_{k \to \infty} H^{(k)} = \frac{J}{n},
  \]
  where $J \in \R^{n \times n}$ is the all-ones matrix.  The \emph{collapse rate} is
  the smallest $k^*$ such that $\mathrm{Var}(H^{(k^*)}) < \tau_{\mathrm{tol}}$ for a
  tolerance $\tau_{\mathrm{tol}} > 0$.
\end{definition}

\begin{theorem}[Sinkhorn Dynamics Collapse]
  \label{thm:sinkhorn_collapses}
  Under the iterated Sinkhorn update $H^{(k+1)} = \mathrm{Sink}_t(H^{(k)})$ starting
  from any doubly stochastic $H^{(0)} \in \mathcal{B}_n$, the sequence converges to
  $J/n$:
  \[
    \left\|H^{(k)} - \frac{J}{n}\right\|_F \;\leq\; C \cdot \lambda_2^k,
  \]
  where $\lambda_2 < 1$ is the second-largest eigenvalue of $H^{(0)}$ and $C$ depends
  only on $H^{(0)}$.  In particular, $\mathrm{Var}(H^{(k)}) \to 0$ exponentially.
\end{theorem}

\begin{proof}[Proof sketch]
  Doubly stochastic matrices have $\lambda_1 = 1$ as the Perron eigenvalue with
  eigenvector $\mathbf{1}/\sqrt{n}$.  Every other eigenvalue satisfies $|\lambda_i| \leq
  \lambda_2 < 1$ (strict inequality for primitive matrices).  Repeated application of
  $\mathrm{Sink}_t$ preserves doubly stochastic structure and contracts all non-Perron
  components by $\lambda_2^t$ per iteration.  The full proof is in
  Appendix~\ref{app:proof_sinkhorn}.
\end{proof}

\subsection{SpectralSphere Manifold}

\begin{definition}[Spectral-Norm Ball]
  For $r > 0$, the spectral-norm ball is
  $\mathcal{S}_r = \{H \in \R^{n \times n} : \spnorm{H} \leq r\}$.
\end{definition}

\begin{definition}[Spectral Projection]
  \label{def:spectral_proj}
  The spectral projection $\Pi_r : \R^{n \times n} \to \mathcal{S}_r$ clips singular
  values to $r$:
  \[
    \Pi_r(H) = U \cdot \mathrm{diag}\!\left(\min(\sigma_i, r)\right) \cdot V^\top,
  \]
  where $H = U \Sigma V^\top$ is the singular value decomposition.
\end{definition}

\begin{definition}[SpectralSphere Update Rule]
  \label{def:spectral_sphere_update}
  The SpectralSphere update with residual strength $\alpha \in (0,1)$ and radius $r > 0$
  is:
  \[
    H^{(k+1)} = (1 - \alpha)\,\Pi_r\!\left(H^{(k)}\right) + \alpha I.
  \]
\end{definition}

The key structural difference from Sinkhorn: $\Pi_r$ projects onto a \emph{ball}
(preserving spectral diversity) rather than an \emph{affine submanifold} (the Birkhoff
polytope, which drives to the centroid $J/n$).  The residual term $\alpha I$ acts as a
leaky integrator, ensuring the identity is always represented in the limit.

\begin{theorem}[SpectralSphere Prevents Collapse]
  \label{thm:spectral_sphere_prevents_collapse}
  Under the SpectralSphere update (Definition~\ref{def:spectral_sphere_update}) with
  $\alpha \in (0,1)$ and $r > 0$, the sequence $\{H^{(k)}\}$ satisfies
  \[
    \mathrm{Var}\!\left(H^{(k)}\right) \;\geq\; V_\infty \;>\; 0 \quad \forall k \geq 0,
  \]
  where $V_\infty = (1-\alpha)^2\,\mathrm{Var}\!\left(\Pi_r(H^*)\right) > 0$ is the
  eigenvalue variance of the fixed point.  In particular, BUC does not occur.
\end{theorem}

\begin{proof}[Proof sketch]
  The fixed point $H^* = (1-\alpha)\Pi_r(H^*) + \alpha I$ has eigenvalue structure
  determined by the residual: eigenvalues of $\alpha I$ are all equal to $\alpha$
  (variance zero), but $\Pi_r(H^*)$ contributes non-trivially unless $r = 0$.  For
  $r > 0$ and generic initializations, the fixed point has $\mathrm{Var}(H^*) > 0$
  (full proof in Appendix~\ref{app:proof_main}).
\end{proof}

\subsection{Neural ODE Formulation}

We lift the discrete SpectralSphere update to continuous time.  Let $\Phi : \R^{n \times
n} \times \R \to \R^{n \times n}$ be a neural network parameterized by $\theta$.  The
trust dynamics ODE is:
\begin{equation}
  \label{eq:trust_ode}
  \frac{dH}{dt} = f_\theta(H, t) := -\nabla_H \mathcal{L}(H) + \alpha (I - H),
\end{equation}
where $\mathcal{L}(H) = \frac{1}{2}\max(0, \spnorm{H} - r)^2$ is the spectral
constraint penalty and $\alpha(I - H)$ is the continuous-time residual injection.

\begin{theorem}[Lyapunov Stability of Trust ODE]
  \label{thm:ode_stability}
  Consider the ODE \eqref{eq:trust_ode} on $\mathcal{S}_r$.  The Lyapunov function
  \[
    V(H) = \frac{1}{2}\left\|H - H^*\right\|_F^2
  \]
  satisfies $\dot{V}(H) \leq -2\alpha V(H)$ along trajectories, establishing global
  exponential stability with rate $2\alpha$.
\end{theorem}

\begin{proof}[Proof sketch]
  Differentiating: $\dot{V} = \langle H - H^*, f_\theta(H, t) \rangle_F$.  The spectral
  penalty term is zero inside $\mathcal{S}_r$.  The residual term contributes
  $-\alpha \langle H - H^*, H - H^* \rangle_F = -\alpha \|H - H^*\|_F^2 = -2\alpha V$.
\end{proof}

\paragraph{ODE stability benchmark.}
The local reproducibility harness reports the Neural ODE's variance retention and
maximum observed spectral norm under projected RK4 integration.  We intentionally do not
claim a Lebesgue-volume topological-capacity result: high topological volume would prove
only that collapse is absent, not that the ODE amplifies the correct specialization.
Routing quality depends on the trained $f_\theta$ and is left to future work, as
discussed in Section~\ref{sec:conclusion}.

\subsection{Differential Privacy via Residual Sensitivity}

\begin{lemma}[$\ell_2$ Sensitivity with Residual Injection]
  \label{lem:residual_sensitivity}
  Let $f(D) = \Pi_r(H_{\mathrm{new}})$ be the spectral projection with $\ell_2$
  sensitivity $\Delta f \leq 2r$ (from the diameter of $\mathcal{S}_r$).  Let
  $g(D) = (1-\alpha)f(D) + \alpha I$ be the residual-injected function.  Then:
  \[
    \left\|g(D) - g(D')\right\|_2
      = (1-\alpha)\left\|f(D) - f(D')\right\|_2
      \leq 2(1-\alpha)r =: \Delta g.
  \]
  The $\alpha I$ terms cancel exactly.  The Gaussian mechanism with noise $Z \sim
  \mathcal{N}(0, \sigma^2 I)$ achieves $(\varepsilon, \delta)$-DP if:
  \[
    \sigma = \frac{2(1-\alpha)r \cdot \sqrt{2 \ln(1.25/\delta)}}{\varepsilon}.
  \]
  At $\alpha = 0.1$, required noise is reduced by exactly $10\%$ relative to the
  baseline $\Delta f = 2r$.
\end{lemma}

\begin{proof}
  $\|g(D) - g(D')\|_2 = \|(1-\alpha)f(D) + \alpha I - (1-\alpha)f(D') - \alpha I\|_2
  = (1-\alpha)\|f(D) - f(D')\|_2 \leq 2(1-\alpha)r$. \qed
\end{proof}

\paragraph{Significance.}
The factor $(1-\alpha)$ means that the same $\alpha$ controlling topological stability
also controls DP noise.  There is no tradeoff: more stability ($\alpha$ larger) implies
less required noise.  At $\alpha = 0.2$ the noise budget shrinks by $20\%$; at $\alpha =
0.5$ by $50\%$.  This synergy is a structural consequence of the algebraic form of
residual injection and holds for any $r > 0$ and any $(\varepsilon, \delta)$.

\begin{algorithm}[t]
  \caption{SpectralSphere Trust Update with DP Broadcast}
  \label{alg:spectral_dp}
  \SetAlgoLined
  \KwIn{Local trust matrix $H_{\mathrm{new}} \in \R^{n \times n}$; radius $r$;
        residual $\alpha$; privacy budget $(\varepsilon, \delta)$}
  \KwOut{Noisy public matrix $\tilde{H}$; noise level $\sigma$}
  \BlankLine
  \tcp{Step 1: Spectral projection}
  $U, \Sigma, V^\top \leftarrow \mathrm{SVD}(H_{\mathrm{new}})$\;
  $H_{\mathrm{proj}} \leftarrow U \cdot \mathrm{diag}(\min(\sigma_i, r)) \cdot V^\top$\;
  \BlankLine
  \tcp{Step 2: Residual injection}
  $H_{\mathrm{proj}} \leftarrow (1-\alpha)\,H_{\mathrm{proj}} + \alpha I$\;
  \BlankLine
  \tcp{Step 3: Calibrate Gaussian noise (Lemma~\ref{lem:residual_sensitivity})}
  $\sigma \leftarrow 2(1-\alpha)r \cdot \sqrt{2 \ln(1.25/\delta)} \,/\, \varepsilon$\;
  $Z \sim \mathcal{N}(0, \sigma^2 I_{n^2})$\;
  \BlankLine
  \tcp{Step 4: DP-noised broadcast matrix}
  $\tilde{H} \leftarrow H_{\mathrm{proj}} + Z$\;
  \Return $(\tilde{H},\, \sigma)$\;
\end{algorithm}
