\section{Related Work}
\label{sec:related}

\paragraph{Multi-agent trust and reputation.}
Early trust models in multi-agent systems \cite{sabater2005review} used scalar reputation
scores aggregated via weighted averages.  More recent work represents trust as probability
distributions \cite{josang2007subjective} or Bayesian networks \cite{teacy2012travos}.
None of these formulations identify or address BUC because they do not use doubly
stochastic normalization; BUC is a property specific to matrix-valued trust systems that
invoke Sinkhorn--Knopp or equivalent row-column normalization.

\paragraph{Sinkhorn--Knopp normalization.}
The Sinkhorn--Knopp algorithm \cite{sinkhorn1964relationship,sinkhorn1967concerning}
converges to the unique doubly stochastic matrix in the same ``scaling class'' as the
input.  It is widely used in optimal transport \cite{cuturi2013sinkhorn}, attention
mechanisms \cite{sinkhorn_transformer}, and — most relevantly — the mHC architecture
\cite{xie2025mhc}, which projects residual connection matrices onto the Birkhoff polytope
to restore identity-mapping properties in deep networks.  Our work identifies the failure
mode latent in all such systems when Sinkhorn is applied \emph{iteratively}: the
long-run fixed point is the uniform matrix, not the intended specialization structure.

\paragraph{Manifold-constrained neural architectures.}
Hyper-Connections \cite{zhu2024hyper} expand the residual stream width in transformers by
replacing the scalar residual coefficient with a learnable matrix $H^{\mathrm{res}}$.
Manifold-Constrained Hyper-Connections (mHC) \cite{xie2025mhc} project $H^{\mathrm{res}}$
onto the Birkhoff polytope, proving bounded spectral norm and compositional closure.
Spectral-Sphere-Constrained Hyper-Connections (sHC) \cite{shc2025spectral} replace the
Birkhoff projection with spectral-norm-ball projection, achieving the same stability
properties with a richer manifold.  Our SpectralSphere Manifold draws directly on sHC's
projection operator, adapting it to the trust dynamics setting.

\paragraph{Neural ODEs.}
Continuous-depth networks \cite{chen2018neural} model hidden-state evolution as the
solution of an ODE $dh/dt = f_\theta(h, t)$.  Augmented Neural ODEs
\cite{dupont2019augmented} extend this to avoid trajectory intersection.  We formulate
trust evolution as a Neural ODE on the spectral sphere, using the ODE framework to prove
global stability rather than as a learned architecture.

\paragraph{Differential privacy for multi-agent systems.}
Federated learning with differential privacy \cite{mcmahan2018dp,geyer2017dp} adds
calibrated Gaussian noise to model updates before aggregation.  The Gaussian mechanism
\cite{dwork2014algorithmic} requires calibration to the $\ell_2$ sensitivity of the
function being privatized.  Our Lemma~\ref{lem:residual_sensitivity} tightens the
sensitivity bound from $2r$ to $2(1-\alpha)r$ via the algebraic cancellation of $\alpha I$
terms, reducing noise requirements by exactly $\alpha \cdot 100\%$.  A related
sensitivity analysis appears in federated Sinkhorn \cite{federated_sinkhorn2025}, but
without the residual-injection tightening.

\paragraph{Constitutional AI and MACI.}
Constitutional AI \cite{bai2022constitutional} provides principles-based governance for
individual LLMs.  Multi-Agent Constitutional Integrity (MACI) extends this to swarms by
enforcing MACI separation of powers: Proposer/Validator/Executor roles are assigned to
distinct agents \cite{acgs2025}.  Our work identifies BUC as a structural threat to MACI:
if trust collapse makes all agents interchangeable, role separation becomes nominal.
SpectralSphere projection restores the specialization diversity that MACI requires.

\paragraph{Lyapunov stability for dynamical systems on manifolds.}
Classical Lyapunov theory \cite{lyapunov1907} establishes global asymptotic stability via
a scalar function $V(x) > 0$ with $\dot{V}(x) < 0$.  LaSalle's invariance principle
\cite{lasalle1960some} extends this to compact manifolds without strict Lyapunov functions.
We apply LaSalle's principle to the spectral sphere ODE, proving that all trajectories
converge to the largest invariant subset of $\{\dot{V} = 0\}$, which excludes the
collapsed uniform matrix.
